The value of being indispensable: a cooperative approach to bargaining with a continuum of players

### 2024-02-22 src/paper/theory.tex
Highlight main theorem
Compare example model in steps

Let us now examine the weighted value in the multi-sided case.
The set of players and the characteristic function are the same as in the previous section, but players have the following weights: player $P$ has weight $\lambda_P$, while players $A^l_i$ have weight $\lambda_l$.
I will show, that as with the Shapley-value, the limit of the weighted value can also be expressed as an integral over a one-dimensional path.
However, this path is not the diagonal of the unit hypercube, and depends on the ratio of the fringe players' weights.

@@

For ease of comparison with the bargaining case, let us also derive the utilities in this case:
\begin{align*}
    u_i^u(n_1, n_2) &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - \frac{\phi_1(u_i)}{\phi'_1(u_i)} \\
    &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - n_i \kappa'_i(n_i).
\end{align*}
Welfare of side $i$ immediately follows in this case, too:
\begin{align*}
    \label{eq:W_u}
    W_i^u(n_1, n_2) = \alpha_i n_i n_j ^ {\gamma_i} + \gamma_j \alpha_j n_j n_i^{\gamma_j} - n_i F_i - n_i^2 \kappa'_i(n_i) - \int_0^{n_i} \kappa_i(s) \ds.
\end{align*}
Clearly, equilibrium utilities must be smaller than in the welfare-maximizing case, due to $\phi_i$ being increasing, the number of entrants is also lower.

@@

Let us now compare the outcomes at the three cases by looking at the entry fees for side $i$.
From the previous results, they are
\begin{align*}
    p_i^* &= \underbracket{F_i}_{(1)} - \underbracket{\alpha_j \gamma_j n_j n_i^{\gamma_j - 1}}_{(2)}, \\
    p_i^u &= \underbracket{F_i}_{(1)} - \underbracket{\alpha_j \gamma_j n_j n_i^{\gamma_j - 1}}_{(2)} + \underbracket{n_i \kappa'_i(n_i)}_{(3)} \\
    p_i^b &= \underbracket{\frac{\lambda_i}{\lambda_P + \lambda_i} F_i}_{(1)} - \underbracket{\frac{\lambda_i \gamma_j}{\lambda_P + \lambda_j + \lambda_i\gamma_j} \alpha_j n_j n_i^{\gamma_j - 1}}_{(2)} \\
    &+ \underbracket{\frac{\lambda_P + \lambda_j\gamma_i}{\lambda_P + \lambda_i + \lambda_j\gamma_i} \alpha_i n_j^{\gamma_i} +
    - \int_0^{n_1} \left[ 1 - \lambda_i \left( \frac{s}{n_i} \right)^{\lambda_P + \lambda_i - 1} \right] \kappa_i(s) \ds}_{(4)},
\end{align*}
for the welfare-maximizing, unilateral price setting and bargaining cases, respectively.

@@

The bargaining case also has the first two factors, albeit they are somewhat different.
As opposed to the first two cases, the price only contains some fraction of the externalities, and the rest is shared with the platform and the other side.
Furthermore, some part of the welfare generated on side $i$ is also shared with the other players, and thus appears in the price (4).
Therefore, bargaining introduces some interesting departures from both the welfare-maximizing and unilateral pricing cases.

Whether these departures lead to a higher or lower total welfare depends on the specific parameters of the model.
For example, if $\lambda_P$ is very high, then $p_i^b > p_i^u > p_i^*$, leading to lower entry and total surplus than not only the welfare-maximizing case, but also the unilateral pricing case.
That is, everybody looses from bargaining.\footnote{
    In such a case, the platform would like to set lower entry fees that what it can achieve via bargaining.
    While in some settings it is very plausible that it can do it, in others commitment problems might prevent it from doing so.
}

@@

Finally, it is always true that $p_1^b < p_1^*$ or $p_2^b < p_2^*$.\footnote{
    Given that $\alpha_i > 0, i \in \{1, 2\}$.
}
The reason is that the platform's profit is always non-negative in the bargaining case, while it is negative in the welfare-maximizing case.
This also implies that the total welfare is strictly lower than in the welfare-maximizing case.
Therefore, while bargaining can be welfare-increasing compared to unilateral price setting, it can not achieve the maximum possible welfare.\footnote{
    One could argue that the welfare maximizing case is not a realistic benchmark anyway, as it assumes that the platform is making a loss.
}

>>>>

Let us now examine the weighted value in the multi-sided case.
The set of players and the characteristic function are the same as in the previous section, but players have the following weights: player $P$ has weight $\lambda_P$, while players $A^l_i$ have weight $\lambda_l$.
This can be considered as the main theorem of the paper, as most of the propositions in the previous sections are special cases of it.

I will show, that as with the Shapley-value, the limit of the weighted value can also be expressed as an integral over a one-dimensional path.
However, this path is not the diagonal of the unit hypercube, and depends on the ratio of the fringe players' weights.

@@

For ease of comparison with the bargaining case, let us also derive the utilities in this case:
\begin{align*}
    u_i^u(n_1, n_2) &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - \frac{\phi_1(u_i)}{\phi'_1(u_i)} \\
    &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - n_i \kappa'_i(n_i).
\end{align*}
Welfare of side $i$ immediately follows in this case, too:
\begin{align*}
    \label{eq:W_u}
    W_i^u(n_1, n_2) = \alpha_i n_i n_j ^ {\gamma_i} + \gamma_j \alpha_j n_j n_i^{\gamma_j} - n_i F_i - n_i^2 \kappa'_i(n_i) - \int_0^{n_i} \kappa_i(s) \ds.
\end{align*}
Clearly, equilibrium utilities must be smaller than in the welfare-maximizing case.
Due to $\phi_i$ being increasing, the number of entrants is also lower.

@@

Let us now compare the outcomes of the three cases described in this section.
To highlight the similarity to \textcite{armstrong2006competition}, I will first consider the case of linear network effects.
Furthermore, for simplicity of exposition, I start by considering the non-weighted Shapley value before moving on to the weighted value.

Assume that network effects are linear, i.e. $\gamma_i = 1$ for $i \in \{1, 2\}$.
Also, focus on non-weighted Shapley values ($\lambda_P = \lambda_1 = \lambda_2 = 1$).
From the previous results, the entry fees for side $i$ are
\begin{align*}
    p_i^* &= \underbracket{F_i}_{(1)} - \underbracket{\alpha_j n_j n_i}_{(2)}, \\
    p_i^u &= \underbracket{F_i}_{(1)} - \underbracket{\alpha_j n_j n_i}_{(2)} + \underbracket{n_i \kappa'_i(n_i)}_{(3)} \\
    p_i^b &= \underbracket{\frac{1}{2} F_i}_{(1)} - \underbracket{\frac{1}{3} \alpha_j n_j n_i}_{(2)} + \underbracket{\frac{2}{3} \alpha_i n_j - \int_0^{n_1} \left[ 1 - \frac{s}{n_i} \right] \kappa_i(s) \ds}_{(4)},
\end{align*}
for the welfare-maximizing, unilateral price setting and bargaining cases, respectively.

@@

The bargaining case also includes the first two factors, albeit they are somewhat different.
As opposed to the first two cases, the price only contains some fraction of the externalities, and the rest is shared with the platform and the other side.\footnote{
    Due to the linearity of the model, various parts of the value are shared equally between players who participate in producing them.
    This is the reason why, for example, the platform and the firms share the cost of serving the entrants (1) equally, or that all three types of players share the value generated on a given side (2) equally.
}
Furthermore, some part of the welfare generated on side $i$ is also shared with the other players, and thus appears in the price (4).
Therefore, bargaining introduces some interesting departures from both the welfare-maximizing and unilateral pricing cases.

Now let us consider the cases of non-linear network effects.
The welfare-maximizing, unilateral pricing and bargaining entry fees are as follows:
\begin{align*}
    p_i^* &= \underbracket{F_i}_{(1)} - \underbracket{\alpha_j \gamma_j n_j n_i^{\gamma_j - 1}}_{(2)}, \\
    p_i^u &= \underbracket{F_i}_{(1)} - \underbracket{\alpha_j \gamma_j n_j n_i^{\gamma_j - 1}}_{(2)} + \underbracket{n_i \kappa'_i(n_i)}_{(3)} \\
    p_i^b &= \underbracket{\frac{1}{2} F_i}_{(1)} - \underbracket{\frac{\gamma_j}{2 + \gamma_j} \alpha_j n_j n_i^{\gamma_j - 1}}_{(2)} + \underbracket{\frac{1 + \gamma_i}{2 + \gamma_i} \alpha_i n_j^{\gamma_i} - \int_0^{n_1} \left[ 1 - \frac{s}{n_i} \right] \kappa_i(s) \ds}_{(4)}.
\end{align*}
For the most part, the intuition is the same as in the linear case.
The welfare-maximizing price is still the cost of serving one additional player (1) and the positive externality imposed on the other side (2).
In addition to those, the unilaterally set entry fee includes a markup related to the elasticity of entry (3).
Finally, in the case of bargaining, these factors do not appear in full, but are shared with the platform and the other side.

There is one important difference in the bargaining case, however.
In the linear case, the terms related to the value generated on each side were shared between the players in a constant proportion, independent of the model parameters.
In the non-linear case, however, these shares do depend on the shape of the network effects.
In particular, when side $i$'s network externality on side $j$ is more convex (e.g. firms on side $i$ are less substitutable), then it receives a larger share of the value generated on the other side.
This is signified by the fact that the coefficient of side $i$'s network externality, $\frac{\gamma_j}{2 + \gamma_j}$, is increasing in $\gamma_j$.
This result shows that the bargaining outcomes are not just a trivial distribution of the value, but also take into account the shape of the network effects.

Finally, for completeness, let us consider the case of weighted values.
The following theorem summarizes the entry fees in this most general case.
\begin{theorem}
    Let $p_i^*$, $p_i^u$ and $p_i^b$ denote the welfare-maximizing, unilateral pricing and bargaining entry fees, respectively.
    Then,
    \begin{align*}
        p_i^* &= \underbracket{F_i}_{(1)} - \underbracket{\alpha_j \gamma_j n_j n_i^{\gamma_j - 1}}_{(2)}, \\
        p_i^u &= \underbracket{F_i}_{(1)} - \underbracket{\alpha_j \gamma_j n_j n_i^{\gamma_j - 1}}_{(2)} + \underbracket{n_i \kappa'_i(n_i)}_{(3)} \\
        p_i^b &= \underbracket{\frac{\lambda_i}{\lambda_P + \lambda_i} F_i}_{(1)} - \underbracket{\frac{\lambda_i \gamma_j}{\lambda_P + \lambda_j + \lambda_i\gamma_j} \alpha_j n_j n_i^{\gamma_j - 1}}_{(2)} \\
        &+ \underbracket{\frac{\lambda_P + \lambda_j\gamma_i}{\lambda_P + \lambda_i + \lambda_j\gamma_i} \alpha_i n_j^{\gamma_i} - \int_0^{n_1} \left[ 1 - \lambda_i \left( \frac{s}{n_i} \right)^{\lambda_P + \lambda_i - 1} \right] \kappa_i(s) \ds}_{(4)}.
    \end{align*}
\end{theorem}

In the case of the linear terms, such as the cost of serving one additional player (1), the addition of bargaining weights simply leads to players shareing the cost in a constant proportion, according to their weights.
The higher the weight, the larger the share for the player.
While the latter is true for non-linear terms, as well, the influence of the weights is more complex, as it also depends on the shape of value function.
For example, in term (2), $\lambda_i$ interacts with the parameter determining the shape of the network effects, $\gamma_j$
Therefore, the weighted value is more than just a simple reweighting of the non-weighted value, and its effects can be quite complex depending on the assumptions on the value function.

Whether bargaining leads to a higher or lower total welfare depends on the parameters of the model.
For example, if $\lambda_P$ is very high, then $p_i^b > p_i^u > p_i^*$, leading to fewer firms entering and lower total surplus than not only the welfare-maximizing case, but also the unilateral pricing case.
That is, everybody looses from bargaining.\footnote{
    In such a case, the platform would like to set lower entry fees that what it can achieve via bargaining.
    While in some settings it is very plausible that it can do it, in others commitment problems might prevent it from doing so.
}

@@

Finally, it is always true that $p_1^b < p_1^*$ or $p_2^b < p_2^*$.
The reason is that the platform's profit is always non-negative in the bargaining case, while it is negative in the welfare-maximizing case.
This also implies that the total welfare is strictly lower than in the welfare-maximizing case.
Therefore, while bargaining can be welfare-increasing compared to unilateral price setting, it can not achieve the highest possible total welfare.\footnote{
    One could argue that the welfare maximizing case is not a realistic benchmark anyway, as it assumes that the platform is making a loss.
}
